---
title: "Reproducible Research: Peer Assessment 1"
author: Ha Nguyen
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r, echo=TRUE}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", "activity.zip", method = "curl")
unzip("activity.zip",files = "activity.csv")
activity <- read.csv("activity.csv", stringsAsFactors = F)
```

## What is mean total number of steps taken per day?
### Total number of steps taken each day 
```{r, echo=TRUE}
library(dplyr)
activity_per_day <- activity %>% group_by(date) %>% summarise(totalSteps = sum(steps, na.rm = T))
activity_per_day
```

### Histogram 
```{r, echo=TRUE}
hist(activity_per_day$totalStep)
```

### Mean and median of the total steps taken per day 
``` {r, echo=TRUE}
summary(activity_per_day$totalStep)
``` 

## What is the average daily activity pattern?
### Get the mean steps per time interval
```{r, echo=TRUE}
activity_per_interval <- activity %>% group_by(interval) %>% summarise(meanSteps = sum(steps, na.rm=T)/length(date))
activity_per_interval
```

### Time series plot of the 5-minute interval 
```{r, echo=TRUE}
with(activity_per_interval, plot(interval, meanSteps, type = "l", main = "Average daily activity")) 
```

* interval from 8-9 am contains the maximum number of steps 

## Imputing missing values
### Number of missing values in the data set 
```{r, echo=TRUE}
sum(is.na(activity$steps ))
```

### Filling all missing values with the mean for that 5-min interval 
```{r, echo=TRUE}
# Create copy of data set and filling missing values in the data set 
activityComp <- activity
# Record interval having missing values 
intervalMissingV <- activityComp$interval[is.na(activityComp$steps)]
# Replace them with mean of 5 min interval corresponded 
activityComp$steps[is.na(activity$steps)]<- activity_per_interval$meanSteps[match(intervalMissingV, activity_per_interval$interval)]
head(activityComp)
```

### Histogram of steps taken per day 
```{r,  echo = T}
activityComp_per_day <- activityComp %>% group_by(date) %>% summarise(totalSteps = sum(steps, na.rm = T))
hist(activityComp_per_day$totalSteps)
```

### Mean and median of total number of steps taken per day with missing values replaced 
```{r,  echo = T}
summary(activityComp_per_day$totalStep)
```

### Compared with Old mean and median of total number of steps taken per day with missing values 
```{r,  echo = T}
summary(activity_per_day$totalStep)
```

- Affect of Imputing missing data on the estimates of the total daily number of steps: 

        +  Changing Mean values
        +  Keeping Median value the same.  

## Are there differences in activity patterns between weekdays and weekends?
### Convert date column into date type 
```{r, echo=TRUE}
activityComp$date <- strptime(activityComp$date, format = "%F")
```

### Find weekday for each date column 
``` {r, echo=TRUE}
activityComp$weekday <- weekdays(activityComp$date)
```

### Create factor variable to identify weekday / weekend 
```{r, echo=TRUE}
weekday_list <- c("Monday", "Tuesday","Wednesday", "Thursday", "Friday")
activityComp$isWeekday <- as.factor(ifelse(activityComp$weekday %in% weekday_list, "weekday", "weekend"))
```

### Time Series Plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days
```{r, echo=TRUE}
library(ggplot2)
activityComp$date <- as.character(activityComp$date)
activityComp_per_weekday <- activityComp %>% group_by(isWeekday, interval) %>% summarise(meanSteps = sum(steps, na.rm=T)/length(isWeekday))
activityComp_per_weekday
qplot(data = activityComp_per_weekday, x = interval, y = meanSteps, 
      geom = "path", facets =isWeekday~.) 
```

- It seems there were more steps/activities in weekday than in weekend. 
